---
title: "lab week 5 time series and a map"
author: "Maya Szafraniec"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE, warning = FALSE)
```

## attach packages

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forcats)
library(sf)
library(tmap)
library(mapview)

```

# us energy

```{r}

us_renew<-read_csv(here("data","renewables_cons_prod.csv")) %>% 
  clean_names()

```


Make the discription all lowercase and only keep options when discription variable only contains one pattern of consuption. remove any obs where discription includes the word total

```{r}
renew_clean<-us_renew %>%
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern ="consumption")) %>% 
  filter(!str_detect(description, pattern ="total"))

```

convert yyyymm to a date 

```{r}

renew_date<-renew_clean %>% 
  mutate(yr_mo_day=lubridate::parse_date_time(yyyymm,"ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value=as.numeric(value)) %>% 
  drop_na(month_sep,value)

#make a verison where month and year are in seperate columns

renew_parsed<-renew_date %>% 
  mutate(month=month(yr_mo_day,label = TRUE)) %>% 
  mutate(year=year(yr_mo_day))


```

# look at it 

```{r}
renew_gg<-ggplot(renew_date, aes(x = month_sep, y=value, group = description))+
  geom_line(aes(color=description)) #you dont have to do both group and color they are redundent

renew_gg

view(palettes_d_names) #look at all the colors in paletter
```

#update color in renew gg

```{r}
renew_gg+scale_color_paletteer_d("calecopal::kelp1")
```

## coerce r renew_parsed to a tsibble

```{r}
renew_ts<-as_tsibble(renew_parsed, key = description, index = month_sep)


```

Lets look at time shit in a few ways
```{r}
renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value)

renew_ts %>% gg_season(value)

ggplot(data=renew_parsed, aes(x=month,y=value,group=year))+
  geom_line(aes(color=year))+
  facet_wrap(~description,
             ncol=1,
             scale="free",
             strip.position="right")


```

just look at hydroelectric 

```{r}
hydro_ts<-renew_ts %>% 
  filter(description=="hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)

ggplot(hydro_ts,aes(x=month,y=value,group=year))+
  geom_line(aes(color=year))


```

### what if i want the quarterly consumption for hydro

```{r}

hydro_quarterly<-hydro_ts %>% 
  index_by(year_qu=~(yearquarter(.))) %>% 
  summarize(avg_consumption=mean(value))

head(hydro_quarterly)

```

# lets decompose that time data 

```{r}
dcmp<-hydro_ts %>% 
  model(STL(value~season(window=5)))


components(dcmp) %>% autoplot

```


```{r}
hist(components(dcmp)$remainder) # check if remainders are normally distributed
```

Now lets look at ACF

```{r}

hydro_ts %>% 
  ACF(value) %>% 
  autoplot()

#observation 12 months apart are more highly corrolated than any other plot 

```

# DANGER DANGER read a god damn book first

```{r}
hydro_model<-hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h="4 years")

hydro_model %>% autoplot(filter(hydro_ts,year(month_sep)>2010))


```


